### 2018-06-11 Writing/response_to_reviewers.tex

\subsubsection{CD as a metric to compare the splits between characters}
Furthermore, the CD metric, used in this manuscript as a proxy for character correlation actually measures the difference between in characters in terms of phylogenetic signal.
The character \texttt{X} described above implies 3 splits for four taxa (and of course, the character \texttt{X'} implies the exact same splits).
Another character \texttt{Y = {1, 1, 1, 2}} will imply only two splits, non shared with \texttt{X} and a character \texttt{Z = {0, 2, 1, 0}} (\texttt{Z' = {1, 2, 3, 1}}) would imply the same splits as \texttt{X} (hence, $CD_{(X',Z')} = 0$ and $CD_{(X',Y')} = 0.5$).

Following these two points, the reviewer counter example $CD_{(x,y)} = 0$ means that 1) $x'$ and $y'$ are identical and the characters $x$ and $y$ imply the same number of splits in a tree (here no splits since only one taxon is used - however, this would hold for any $n$ number of taxa).
We have clarified these two points in the manuscript and in the mathematical demonstration.

%TG: TODO: introduce this

\textit{(i.e. entirely correlated characters would define identical splits between a set of taxa)}. lines @@@

\textit{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.} lines @@@

### 2018-06-11 Writing/draft_highlighted.tex
Added precisions for the CD metric

\subsection{Measuring differences between characters}
\label{CDdescription}
To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information.
In order to measure this, we propose a new distance metric to measure the difference between two characters:

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
\hl{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
\citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

>>>>

\subsection{Measuring differences between characters}
\label{CDdescription}
To measure the effect of character correlation as interpreted by the phylogenetic software, we define characters as being entirely correlated if they give the same phylogenetic information \hl{(i.e. entirely correlated characters would define identical splits between a set of taxa)}.
In order to measure this, we propose a new distance metric to measure the difference between two characters:

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by order of appearance.
In other words, we replaced all the occurrences of the first token to be $1$, the second to be $2$, etc.
\hl{This procedure allows to compare characters regardless of the significance given to their tokens (following the \textit{xyz} notation in }
\citealt{felsenstein2004inferring}; \hl{as used in} \citealt{Davalos01072014}).
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}}.
\hl{Similarly, an hypothetical comparison of two character for only one taxa \texttt{x = \{0\}} and \texttt{y = \{1\}} would result in comparing two identical characters \texttt{x' = \{1\}} and \texttt{y' = \{1\}}.}
Note that in terms of phylogenetic signal, both \texttt{A} and \texttt{A'} are exactly identical (forming three distinct splits in the tree inference process).

### 2018-06-11 Writing/supplementary/Sup1-CharDiff.tex
Added precisions for the CD metric

$CD$, the Character Difference metric, is a distance that measures the difference between two characters in terms of shared phylogenetic splits.
For two characters $x$ and $y$, $CD=0$ implies that the characters are identical, i.e. they convey the same phylogenetic signal in term of splits; $CD=1$ implies that the characters are the most different, i.e. they convey the most different phylogenetic signal in terms of splits.

@@

>>>>

$CD$, the Character Difference metric, is a distance that measures the difference between two characters in terms of shared phylogenetic splits.

For two characters $x$ and $y$, $CD=0$ implies that the characters are identical, i.e. they convey the same phylogenetic signal in term of splits; $CD=1$ implies that the characters are the most different, i.e. they convey the most different phylogenetic signal in terms of splits.

@@

\subsection{Justification for the character states translation}
Because the character states tokens are not representing any meaningful numerical value if the characters are unordered (e.g. there is no assumption that state 0 should come before state 1), they need to be reordered in a standard way so that the first token in the character is renamed to ``1'', the second to ``2'' etc.
For example, in a 3 state character for four taxa: \texttt{X = {1, 0, 2, 1}}, the character tokens that are integers representing the states could equally have been \texttt{X = {E, A, I, E}} or \texttt{X = {blue, red, absent, blue}}.
Therefore, to compare characters fairly, the tokens needs to be translated in a standard way such as the first token is replaced by the symbol ``1'', etc.
In this example \texttt{X = {1, 0, 2, 1}} becomes \texttt{X' = {1, 2, 3, 1}}

\subsection{Interpretation}
For a characters \texttt{X = {1, 0, 2, 1}} implying three splits between four taxa and two characters \texttt{Y = {1, 1, 1, 2}} and \texttt{Z = {0, 2, 1, 0}} implying respectively two and three splits between four taxa, $CD_{(X,Y)} = 0.5$ and $CD_{(X,Z)} = 0$.
This can be interpreted as the probability of characters \texttt{X}, \texttt{Y} and \texttt{X}, \texttt{Z} giving a conflicting phylogenetic signal between the four taxa.
Here \texttt{X}, \texttt{Y} will conflict for 50\% of the taxa and \texttt{X}, \texttt{X} for 0\% of them.
